---
title: "Merge of Political Data"
output: html_document
date: "2024-04-16"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

knitr::opts_knit$set(root.dir="C:/Users/giorg/OneDrive/Documenti/RA Silvia Pianta/Political Parties Climate Discourse")

```

```{r libraries}

library(dplyr)
library(tidyr)
library(haven)
library(readr)
library(lubridate)
library(countrycode)
library(expss)

```

This code joins together:

-   data on climate policy stringency and the number of climate policies implemented in different sectors, from the OECD's Climate Actions and Policies Measurement Framework (CAPMF, publicly available [at this link](https://data-explorer.oecd.org/vis?lc=en&df[ds]=DisseminateFinalDMZ&df[id]=DSD_CAPMF%40DF_CAPMF&df[ag]=OECD.ENV.EPI&df[vs]=1.0&av=true&pd=%2C&dq=.A...&to[TIME_PERIOD]=false&ly[rw]=REF_AREA&ly[cl]=TIME_PERIOD&vw=ov));

-   data on elected governments and parliaments from WZB's Political Parties, Presidents, Elections and Governments database (PPEG, publicly available [at this link](https://ppeg.wzb.eu/), combined version).

### OECD CAPMF

```{r oecd_capmf_data}

# Import the data
oecd_capmf <- read_csv("Dati/OECD.ENV.EPI,DSD_CAPMF@DF_CAPMF,1.0+all.csv")

# Target countries
countries <- c("Austria", "Belgium", "Bulgaria", "Croatia", "Czechia",
               "Denmark", "Estonia", "Finland", "France", "Germany", "Greece",
               "Hungary", "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
               "Malta", "Netherlands", "Poland", "Portugal", "Romania",
               "Slovak Republic", "Slovenia", "Spain", "Sweden", "United Kingdom")

# Target policies (see documentation for description)
policies <- c("LEV1_SEC", "LEV2_CROSS_SEC_RDD", "LEV2_CROSS_SEC_FFPP")

# Clean
oecd_capmf %>%
  relocate(REF_AREA, `Reference area`, TIME_PERIOD, CLIM_ACT_POL,
           `Climate actions and policies`, MEASURE, Measure) %>%
  relocate(STRUCTURE, STRUCTURE_ID, STRUCTURE_NAME, .after = last_col()) %>%
  arrange(REF_AREA, TIME_PERIOD, CLIM_ACT_POL, MEASURE) %>%
  # Select type of policy
  filter(CLIM_ACT_POL %in% policies) %>% 
  # Select geographical scope
  filter(`Reference area` %in% countries) %>%
  # Keep relevant variables
  select(!c("Climate actions and policies", "Measure", "ACTION",
            "FREQ", "Frequency of observation", "UNIT_MEASURE",
            "Unit of measure", "Time period", "Observation value",
            "Observation status", "UNIT_MULT", "Unit multiplier",
            "DECIMALS", "Decimals", "STRUCTURE", "STRUCTURE_ID",
            "STRUCTURE_NAME")) %>%
  # Pivot wider
  rename(OS = OBS_STATUS) %>%
  pivot_wider(names_from = c(CLIM_ACT_POL, MEASURE),
              names_glue = "{CLIM_ACT_POL}_{MEASURE}_{.value}",
              values_from = c(OBS_VALUE, OS)) -> oecd_capmf_clean

# Discard remaining objects
rm(countries, policies)

```

### Elected governments data

```{r wzb_ppeg_data}

### Import the data

# Read .dta file
ppeg_comb <- read_dta("Dati/ppeg_comb_2024v1/ppeg_comb_2024v1.dta") 

# Target countries
countries <- c("Austria", "Belgium", "Bulgaria", "Croatia", "Czech Republic",
               "Denmark", "Estonia", "Finland", "France", "Germany", "Greece",
               "Hungary", "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
               "Malta", "Netherlands", "Poland", "Portugal", "Romania",
               "Slovakia", "Slovenia", "Spain", "Sweden", "United Kingdom")


### Clean and obtain additional information

# Create columns with year of the election dates
ppeg_comb %>%
  mutate(eyear = year(edate), .after = edate) -> ppeg_comb

# Keep only relevant countries
ppeg_comb %>% filter(cname_en %in% countries) -> ppeg_comb

# Create new column with the duration of each government (time elapsed between
# government inaugurations)
ppeg_comb %>% distinct(cname_en, idate) %>%
  group_by(cname_en) %>%
  arrange(idate, .by_group = T) %>%
  mutate(previous_gov_idate = lag(idate),
         gov_duration = (lead(idate) - idate)) -> gov_duration
ppeg_comb %>% left_join(gov_duration, by = c("cname_en", "idate")) %>%
  relocate(previous_gov_idate, gov_duration, .after = idate) -> ppeg_comb

# Create new column with the duration of each parliament (time elapsed between
# each parliamentary election
ppeg_comb %>% distinct(cname_en, edate) %>%
  group_by(cname_en) %>% arrange(edate, .by_group = T) %>%
  mutate(previous_ele_edate = lag(edate),
         ele_duration = (lead(edate) - edate)) -> ele_duration
ppeg_comb %>% left_join(ele_duration, by = c("cname_en", "edate")) %>%
  relocate(previous_ele_edate, ele_duration, .after = eyear) -> ppeg_comb

# Discard remaining objects
rm(ele_duration, gov_duration, countries)


### Expand the dataset to include rows for every year in between the
### inaugurations of elected governments

# Expand
ppeg_comb %>% 
  mutate(iyear = year(idate)) %>% # Column with year of the inauguration date
  select(iso3c, idate, iyear) %>%
  distinct() %>%
  group_by(iso3c) %>%
  complete(iyear = seq(min(iyear), max(iyear), by = 1)) %>%
  fill(idate, .direction = "down") -> ppeg_comb_expanded

# Include the information on parties, election years, etc. (m:m merge) and clean
merge(ppeg_comb_expanded, ppeg_comb,
      by = c("iso3c", "idate"),
      all.x = T, all.y = T) %>%
  relocate(idate, iyear, .before = previous_gov_idate) %>%
  mutate(TIME_PERIOD = iyear, .after = cname_en) %>%
  mutate(iyear = year(idate)) -> ppeg_comb_clean

```

### Merge and export

Join the data on policy stringency etc. with the data on elected Parliaments/Governments

```{r capmf_ppeg_merge}

# # Check countries correspondence
# ppeg_comb %>%
#   distinct(iso3c, cname_en) -> countries_ppeg
# oecd_capmf_clean %>%
#   distinct(REF_AREA, `Reference area`) -> countries_campf

# Prepare PPEG data for the merge
ppeg_comb_clean %>%
  select(!cname_en) %>%
  rename(REF_AREA = iso3c) -> for_merge

# Merge
left_join(oecd_capmf_clean, for_merge, by = c("REF_AREA", "TIME_PERIOD")) %>%
  relocate(starts_with(c("LEV1", "LEV2")), .after = last_col()) -> capmf_ppeg

# # (OPTIONAL) Create a variable with the chronological order of the government
# # observation in each country
# capmf_ppeg_new <- capmf_ppeg[FALSE, ]
# for (country in unique(capmf_ppeg$REF_AREA)) {
#   capmf_ppeg %>%
#     filter(REF_AREA == country) %>%
#     group_by(REF_AREA, idate) %>%
#     mutate(gov_number = cur_group_id(), .after = iyear) -> capmf_ppeg_filt
#   rbind(capmf_ppeg_new, capmf_ppeg_filt) -> capmf_ppeg_new
# }
# capmf_ppeg_new -> capmf_ppeg
# rm(capmf_ppeg_filt, capmf_ppeg_new)

# Compute the change in policy stringency and in the number of climate policies
# in place at each consecutive government
capmf_ppeg %>%
  # Keep election years only and remove repeated values
  filter(TIME_PERIOD == iyear) %>%
  select(REF_AREA, idate, iyear, ends_with("OBS_VALUE")) %>% 
  distinct() %>%
  # Compute the change
  group_by(REF_AREA) %>% arrange(idate, .by_group = T) %>%
  mutate(across(.cols = ends_with("OBS_VALUE"),
                .fns = ~ lead(.x) - .x,
                .names = "d_{.col}"),
         across(.cols = ends_with("OBS_VALUE") & !starts_with("d_"),
                .fns = ~ (lead(.x) - .x)/.x,
                .names = "pc_{.col}"),
         across(.cols = ends_with("OBS_VALUE") & !starts_with(c("d_", "pc_")),
                .fns = ~ log(.x + 1, base = exp(1)),
                .names = "ln_{.col}")) %>%
  select(REF_AREA, idate, starts_with(c("d_", "pc_", "ln_"))) -> pol_change
# Merge
left_join(capmf_ppeg, pol_change, by = c("REF_AREA", "idate")) -> capmf_ppeg
  

# Add dummy for countries who are also in the WZB's presidential elections
# database
pres_countries <- c("Austria", "Bulgaria", "Croatia", "Czech Republic",
                    "Finland", "France", "Ireland", "Lithuania", "Poland",
                    "Portugal", "Romania", "Slovakia", "Slovenia")
capmf_ppeg %>% 
  mutate(pres_dummy = case_when(`Reference area` %in% pres_countries ~ 1,
                                .default = 0)) -> capmf_ppeg

```

Export in Stata format

```{r stata_export}

# Clean
capmf_ppeg %>%
  # Change variable names
  rename(iso3c = REF_AREA,
         country_name = `Reference area`,
         time_period = TIME_PERIOD) %>%
  rename_with(.cols = ends_with("OBS_VALUE"),
              .fn = ~ substr(.x, 1, (nchar(.x) - nchar("_OBS_VALUE")))) %>%
  rename_with(.cols = starts_with(c("LEV", "d_LEV")),
              .fn = ~ gsub("STRINGENCY", "string", .x)) %>%
  rename_with(.cols = starts_with("LEV"),
              .fn = ~ tolower(substr(.x, nchar("LEV1_")+1, nchar(.x)))) %>% 
  rename_with(.cols = starts_with(c("d_LEV1", "pc_LEV1", "ln_LEV1")),
              .fn = ~ tolower(gsub("LEV1_", "", .x))) %>%
  rename_with(.cols = starts_with(c("d_LEV2", "pc_LEV2", "ln_LEV2")),
              .fn = ~ tolower(gsub("LEV2_", "", .x))) %>%
  
  # Add labels
  apply_labels(iso3c = var_lab(ppeg_comb$iso3c),
               country_name = "Country name",
               time_period = "Year",
               iso2c = var_lab(ppeg_comb$iso2c),
               
               # Variables on parliaments and governments (PPEG)
               edate = var_lab(ppeg_comb$edate),
               eyear = "Year of national lower house election",
               previous_ele_edate = "Date of previous national lower house election",
               ele_duration = "Duration of current national lower house",
               idate = var_lab(ppeg_comb$idate),
               iyear = "Year of inauguration of the Prime Minister/Head of State",
               previous_gov_idate = "Date of previous inauguration of the PM/HS",
               gov_duration = "Duration of current government",
               pm = var_lab(ppeg_comb$pm),
               party_id = var_lab(ppeg_comb$party_id),
               cmp = var_lab(ppeg_comb$cmp),
               cmp_parfam = var_lab(ppeg_comb$cmp_parfam),
               pinitials = var_lab(ppeg_comb$pinitials),
               pname_en = var_lab(ppeg_comb$pname_en),
               pname_or = var_lab(ppeg_comb$pname_or),
               nmin = var_lab(ppeg_comb$nmin),
               nminpp = var_lab(ppeg_comb$nminpp),
               v_share_wgt = var_lab(ppeg_comb$v_share_wgt),
               s_share = var_lab(ppeg_comb$s_share),
               share_remark = var_lab(ppeg_comb$share_remark),
               gov_seat = var_lab(ppeg_comb$gov_seat),
               gov_seat_remark = var_lab(ppeg_comb$gov_seat_remark),
               ngovparties = var_lab(ppeg_comb$ngovparties),
               ngovparties_remark = var_lab(ppeg_comb$ngovparties_remark),
               region = var_lab(ppeg_comb$region),
               continent = var_lab(ppeg_comb$continent),
               eu_since = var_lab(ppeg_comb$eu_since),
               eu_exit = var_lab(ppeg_comb$eu_exit),
               oecd_since = var_lab(ppeg_comb$oecd_since),
               
               # Policy stringency and count variables (OECD)
               sec_pol_count = "Policy count - Sectoral policies",           
               sec_pol_string = "Policy stringency - Sectoral policies",
               sec_pol_count_os = "Policy count - Sectoral policies (obs. status)",
               sec_pol_string_os = "Policy stringency - Sectoral policies (obs. status)",
               cross_sec_ffpp_pol_count = "Policy count - Fossil fuel policies",
               cross_sec_ffpp_pol_string = "Policy stringency - Fossil fuel policies",
               cross_sec_rdd_pol_count = "Policy count - RD&D policies", 
               cross_sec_rdd_pol_string = "Policy stringency - RD&D policies",
               cross_sec_ffpp_pol_count_os = "Policy count - Fossil fuel policies (obs. status)",
               cross_sec_ffpp_pol_string_os = "Policy stringency - Fossil fuel policies (obs. status)",
               cross_sec_rdd_pol_count_os = "Policy count - RD&D policies (obs. status)",
               cross_sec_rdd_pol_string_os = "Policy stringency - RD&D policies (obs. status)",
               d_sec_pol_count = "Change in number of sectoral policies",
               d_sec_pol_string = "Change in policy stringency of sectoral policies",
               d_cross_sec_ffpp_pol_count = "Change in number of fossil fuel policies",
               d_cross_sec_ffpp_pol_string = "Change in stringency of fossil fuel policies",
               d_cross_sec_rdd_pol_count = "Change in number of RD&D policies",
               d_cross_sec_rdd_pol_string = "Change in stringency of RD&D policies",
               pc_sec_pol_count = "Percentage change in number of sectoral policies",
               pc_sec_pol_string = "Percentage change in policy stringency of sectoral policies",
               pc_cross_sec_ffpp_pol_count = "Percentage change in number of fossil fuel policies",
               pc_cross_sec_ffpp_pol_string = "Percentage change in stringency of fossil fuel policies",
               pc_cross_sec_rdd_pol_count = "Percentage change in number of RD&D policies",
               pc_cross_sec_rdd_pol_string = "Percentage change in stringency of RD&D policies",
               ln_sec_pol_count = "Log number of sectoral policies",
               ln_sec_pol_string = "Log policy stringency of sectoral policies",
               ln_cross_sec_ffpp_pol_count = "Log number of fossil fuel policies",
               ln_cross_sec_ffpp_pol_string = "Log stringency of fossil fuel policies",
               ln_cross_sec_rdd_pol_count = "Log number of RD&D policies",
               ln_cross_sec_rdd_pol_string = "Log stringency of RD&D policies",

               pres_dummy = "Dummy for country in WZB's presidential elections database") -> capmf_ppeg


# Export
write_dta(capmf_ppeg, "merged_ppeg_capmf.dta")

```
